Stage-Splitting a Modal Language

### 2014-08-26 paper/semantics.tex
Added tech report.  Changes to semantics section.

Thus, our dynamic semantics for \lang\ evaluates all of a term's stage \bbone\
subexpressions before any of its stage \bbtwo\ subexpressions. This results in a
stage \bbtwo\ term with no stage \bbone\ content; we say this is a term in a
monostaged%
\footnote{In general, \lang could have more than two stages, in which case the
expression would be in a language with one fewer stage.}
language called \langTwo. Then, we perform the remainder of the evaluation with 
$\tworedsym$, an ordinary dynamic semantics for \langTwo.

@@

\subsection{Metatheory}

Recall that residuals live in \langTwo; we will indicate typing judgments in
\langTwo\ with $\vdash_\bbtwo$.

%\begin{definition}
%Context $\Gamma$ is well-formed ($\Gamma\wf$) if it
%contains only stage-2 variables.
%\end{definition}

\begin{definition}
An environment $\xi$ is well-formed ($\Gamma\vdash\xi\wf$) if either:
\begin{enumerate}
\item $\xi = \cdot$; or
\item $\xi = \xi',x:B\mapsto e$ where
$\Gamma\vdash\xi'\wf$ and
$\typeslangTwo[\Gamma,\dom{\xi'}] e B$
%$\Gamma,\dom{\xi'}\vdash \coltwo{e}{B}$ and
%$\Gamma,\dom{\xi'}\vdash e \res$.
\end{enumerate}
\end{definition}

\begin{theorem}
If $\typeswor e A$ then $\Gamma\wf$ and $A\istypewor$.
\end{theorem}

\begin{theorem}
If $\diaonesub$ and $\typesone e A$ then
\begin{enumerate}
\item $\Gamma\vdash\xi\wf$;
\item $\Gamma,\dom\xi\vdash \colone{v}{A}$; and
\item $\Gamma,\dom\xi\vdash v\pval$.
\end{enumerate}
\end{theorem}

\begin{theorem}
If $\diatwosub$ and $\typestwo e A$ then
\begin{enumerate}
\item $\typeslangTwo q A$; and
\item $\Gamma\vdash_\bbtwo q\val$.
\end{enumerate}
\end{theorem}

\begin{theorem}\label{thm:reify-type}
If $\Gamma\vdash\xi\wf$ and
$\Gamma,\dom\xi\vdash \colone{\next\ \hat y}{\fut A}$
then 
$\reify{\xi}{\hat y}{q}$ and
$\typeslangTwo q A$.
\end{theorem}

>>>>

Thus, our dynamic semantics for \lang\ evaluates all of a term's stage \bbone\
subexpressions before any of its stage \bbtwo\ subexpressions. This results in a
stage \bbtwo\ term with no stage \bbone\ content; we say this is a term in a
monostaged
\footnote{In general, \lang could have more than two stages, in which case the
expression would be in a language with one fewer stage.}
language called \langTwo. Then, we perform the remainder of the evaluation with 
$\tworedsym$, an ordinary dynamic semantics for \langTwo.

@@

%\subsection{Metatheory}
%
%Recall that residuals live in \langTwo; we will indicate typing judgments in
%\langTwo\ with $\vdash_\bbtwo$.
%
%%\begin{definition}
%%Context $\Gamma$ is well-formed ($\Gamma\wf$) if it
%%contains only stage-2 variables.
%%\end{definition}
%
%\begin{definition}
%An environment $\xi$ is well-formed ($\Gamma\vdash\xi\wf$) if either:
%\begin{enumerate}
%\item $\xi = \cdot$; or
%\item $\xi = \xi',x:B\mapsto e$ where
%$\Gamma\vdash\xi'\wf$ and
%$\typeslangTwo[\Gamma,\dom{\xi'}] e B$
%%$\Gamma,\dom{\xi'}\vdash \coltwo{e}{B}$ and
%%$\Gamma,\dom{\xi'}\vdash e \res$.
%\end{enumerate}
%\end{definition}
%
%\begin{theorem}
%If $\typeswor e A$ then $\Gamma\wf$ and $A\istypewor$.
%\end{theorem}
%
%\begin{theorem}
%If $\diaonesub$ and $\typesone e A$ then
%\begin{enumerate}
%\item $\Gamma\vdash\xi\wf$;
%\item $\Gamma,\dom\xi\vdash \colone{v}{A}$; and
%\item $\Gamma,\dom\xi\vdash v\pval$.
%\end{enumerate}
%\end{theorem}
%
%\begin{theorem}
%If $\diatwosub$ and $\typestwo e A$ then
%\begin{enumerate}
%\item $\typeslangTwo q A$; and
%\item $\Gamma\vdash_\bbtwo q\val$.
%\end{enumerate}
%\end{theorem}
%
%\begin{theorem}\label{thm:reify-type}
%If $\Gamma\vdash\xi\wf$ and
%$\Gamma,\dom\xi\vdash \colone{\next\ \hat y}{\fut A}$
%then 
%$\reify{\xi}{\hat y}{q}$ and
%$\typeslangTwo q A$.
%\end{theorem}

### 2014-08-26 paper/splitting.tex
Added tech report.  Changes to semantics section.

\section{Splitting Algorithm}

@@

\section{Examples for Staged Pipelines}

Give the gist of one-to-one pipeline example (like client/server).
Then talk about a one-to-many pipeline.
Then talk about a many-to-one pipeline like spark.  It clear how to target something like this for known base types on the boundary, and for product types.  But sums on the boundary are hard!  We leave many-to-one as future-work.

>>>>

\section{Splitting Algorithm}
\label{sec:splitting}

@@

### 2014-08-26 paper/techreport.tex
Added tech report.  Changes to semantics section.

\input{intro}

\input{semantics}

\vfill\eject\pagebreak
\input{splitting}
\input{algoexamples}
\input{relatedwork}

>>>>

\section{Introduction}

This document is a tech report on current research into a program transformation
technique which we call {\em splitting}.  This report has three parts, with
a section dedicated to each,
\begin{itemize}
\item \Cref{sec:semantics} describes a language, called \lang, which unambiguously specifies the inputs to 
our splitting transformation.  It includes both a type system, which is unimaginatively based on previous work, 
and a semantics, which is more novel.
\item \Cref{sec:splitting} describes the splitting transformation itself.  This takes programs, specified in \lang, and converts them to terms in a standard unstaged language.  This section also describes the more formal definition of the correctness for this process.
\item \Cref{sec:examples} gives examples of how this splitting transformation can be used
to derive data structures, and discusses the effectiveness and limitations of this technique.
\end{itemize}

\input{semantics}
\input{splitting}
\input{algoexamples}
